Characterisation of MSC in clinical trial reports: analysis of published descriptors

### 2020-08-10 paper2_figures.Rmd
fig updates for data update

---
title: 'Characterisation of MSC in clinical trial reports: analysis of published descriptors'
author: "Emma Rand"
date: "16/01/2020"
output:
  html_document: default
  svg_document: default
---

@@

fig1 <- article_funnel + data_collection + 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A")&
    theme(plot.tag = element_text(size = 9, hjust = 1, face = "bold"))

ggsave("figs/fig_1.svg", 
       plot = fig1, 
       device = "svg",
       width = fig_w, 
       height = fig_h*1.6,
       units = units)

@@

# by phase
phase_bar <- summary_phase %>% 
  ggplot(aes(x = reorder(phase,n), y = n)) +
  geom_bar(stat = "identity", 
           # fill = pal[4],
           fill = "black",
           size = 0, 
           width = 0.8) +
  xlab("") +
 # geom_text(aes(y = n + 4, label = pc) ) +
  scale_y_continuous(name = NULL, 
                     breaks = seq(0, 35, 10),
                     minor_breaks = seq(0, 35, 1),
                     limits = c(0, 35),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

@@

ggsave("figs/fig_2_black.svg", 
       plot = fig2, 
       device = "svg",
       width = fig_w_2, 
       height = fig_h*2,
       units = units)

@@

fig3 <- source_bar  + allo_auto_bar+ stem_stromal_bar +
  plot_layout(design = layout) +
  plot_annotation(caption = "Source",
                  tag_levels = "A")&
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"),
          plot.caption = element_text(size = 9, hjust = 0.5, vjust = 1, face = "plain"))

ggsave("figs/fig_3_black.svg", 
       plot = fig3, 
       device = "svg",
       width = fig_w_1.5, 
       height = fig_h,
       units = units)
```

@@

### whole fig 4
```{r fig_4-layout}
layout <- "
#aaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaaa
bbbcccccccccccccccccccccccccccccccccccc
"

@@

fig4 <- test_bar + phase_test + markerp_bar +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A")&
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"))

ggsave("figs/fig_4.svg", 
       plot = fig4, 
       device = "svg",
       width = fig_w_2, 
       height = 2.2,
       units = units)
```

@@

fig5 <- fig5a +fig5b &
  plot_layout(design = layout) &
  plot_annotation(tag_levels = "A")&
  geom_hline(yintercept = 1.5) &
   geom_hline(yintercept = 11.5) &
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"),
          plot.margin = margin(0,0,0,0))

ggsave("figs/fig_5.svg", 
       plot = fig5, 
       device = "svg",
       width = fig_w_2, 
       height = fig_h * 1.2,
       units = units)
```

@@

# panel b
fig6b <- moa_isct_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = stem_stromal)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal, 
                    name = "Nomenclature",
                    guide = guide_legend(nrow = 4,
                                         title.position = "top",
                                         title.hjust = 0 )) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 40, 5),
                     minor_breaks = seq(0, 40, 5),
                     limits = c(0, 42),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "MOA") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        plot.margin = margin(t = 0, r = 0, l = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top",
        legend.text = element_text(size = 8, colour = "black"),
        legend.key.size = unit(0.1, "in")) +
  coord_flip()

# panel c
fig6c <- diffmoasum %>%
  ggplot(aes(x = MOA, y = n, fill = differentiation)) +
  geom_bar(stat = "identity", size = 0) +
   scale_fill_manual(values = c("#BABABA", pal[5]), 
                    name = "Differentiation capacity",
                    labels = c("Not demonstrated", "Demonstrated"),
                    guide = guide_legend(nrow = 4, 
                                         title.position = "top",
                                         title.hjust = 0)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 40, 5),
                     minor_breaks = seq(0, 40, 5),
                     limits = c(0, 42),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = NULL, labels =NULL) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        plot.margin = margin(t = 0, r = 0, l = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top",
        legend.text = element_text(size = 8, colour = "black"),
        legend.key.size = unit(0.1, "in")) +
  coord_flip()

@@

fig6 <- fig6a + fig6b + fig6c +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A") &
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"))
ggsave("figs/fig_6.svg", 
       plot = fig6, 
       device = "svg",
       width = fig_w_2, 
       height = 3,
       units = units)

@@

@@

## Figure 10 assocation between differentiation2 and MOA

```{r}
res <- table(dat$differentiation_2, dat$MOA) %>% fisher.test(simulate.p.value = TRUE)
chisq.test(table(dat$differentiation_2, dat$MOA))$exp;table(dat$differentiation_2, dat$MOA)
test_anno <- paste0("There is  significant association between MOA and demonstration of differentiation capacity(", res$method, ": p = ", round(res$p.value, 3),") with more studies demonstrating differentiation for Immune MOA, and fewer for Paracrine and Multiple MOA than expected.")
```
# CD values figures

@@

ggsave("figs/fig_x3_cd_refs_bar.svg", 
       plot = cd_refs_bar, 
       device = "svg",
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

@@

```{r}
cd_values_point_isct <- cd_dat_summary_values %>% 
  filter(ISCT == "ISCT phenotype marker") %>% 
  ggplot(aes(x = marker, y = mean_value)) +
  geom_pointrange(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                  position = position_jitter(width = 0.5), shape = 20) +
  xlab("") +
  scale_y_continuous(name = "Value reported", 
                     breaks = seq(0, 110, 10),
                     minor_breaks = seq(0, 110, 10),
                     limits = c(0, 110),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure X4. TOP" ) +
  geom_vline(xintercept = seq(1.5, 10.5, 1), linetype = "longdash") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4))
ggsave("figs/fig_x4_cd_values_isct.svg", 
       plot = cd_values_point_isct, 
       device = "svg",
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

@@

```{r}
cd_values_point_other <- cd_dat_summary_values %>% 
  filter(ISCT == "Other marker") %>% 
  ggplot(aes(x = marker, y = mean_value)) +
  geom_pointrange(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                  position = position_jitter(width = 0.5), shape = 20) +
  xlab("") +
  scale_y_continuous(name = "Value reported", 
                     breaks = seq(0, 110, 10),
                     minor_breaks = seq(0, 110, 10),
                     limits = c(0, 110),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure X4. TOP" ) +
  geom_vline(xintercept = seq(1.5, 10.5, 1), linetype = "longdash") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4))
ggsave("figs/fig_x5_cd_values_isct.svg", 
       plot = cd_values_point_other, 
       device = "svg",
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

>>>>

---
title: 'Characterisation of MSC in clinical trial reports: analysis of published descriptors'
author: "Emma Rand"
date: "16/01/2020"
output:
  html_document: default
  wmf_document: default
---

@@

fig1 <- article_funnel + data_collection + 
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A") &
    theme(plot.tag = element_text(size = 9, hjust = 1, face = "bold"))

ggsave("figs/fig_1.wmf", 
       plot = fig1, 
       device = "wmf",
       width = fig_w, 
       height = fig_h*1.6,
       units = units)

@@

# by phase
phase_bar <- summary_phase %>% 
  ggplot(aes(x = reorder(phase,n), y = n)) +
  geom_bar(stat = "identity", 
           # fill = pal[4],
           fill = "black",
           size = 0, 
           width = 0.8) +
  xlab("") +
 # geom_text(aes(y = n + 4, label = pc) ) +
  scale_y_continuous(name = NULL, 
                     breaks = seq(0, 20, 10),
                     minor_breaks = seq(0, 20, 1),
                     limits = c(0, 25),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_text(hjust = 0, size = 4)) +
  coord_flip()

@@

ggsave("figs/fig_2_black.wmf", 
       plot = fig2, 
       device = "wmf",
       width = fig_w_2, 
       height = fig_h*2,
       units = units)

@@

fig3 <- source_bar  + allo_auto_bar + stem_stromal_bar +
  plot_layout(design = layout) +
  plot_annotation(caption = "Source",
                  tag_levels = "A") &
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"),
          plot.caption = element_text(size = 9, hjust = 0.5, vjust = 1, face = "plain"))

ggsave("figs/fig_3_black.wmf", 
       plot = fig3, 
       device = "wmf",
       width = fig_w_1.5, 
       height = fig_h,
       units = units)
```

@@

### whole fig 4
```{r fig_4-layout}

@@

fig4 <- test_bar + phase_test + markerp_bar +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A") &
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"))

ggsave("figs/fig_4.wmf", 
       plot = fig4, 
       device = "wmf",
       width = fig_w_2, 
       height = 2.2,
       units = units)
```

@@

fig5 <- fig5a + fig5b &
  plot_layout(design = layout) &
  plot_annotation(tag_levels = "A") &
  geom_hline(yintercept = 1.5) &
   geom_hline(yintercept = 11.5) &
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"),
          plot.margin = margin(0,0,0,0))

ggsave("figs/fig_5.wmf", 
       plot = fig5, 
       device = "wmf",
       width = fig_w_2, 
       height = fig_h * 1.2,
       units = units)
```

@@

# panel b
fig6b <- moa_isct_summary %>% 
  ggplot(aes(x = MOA, y = n, fill = stem_stromal)) +
  geom_bar(stat = "identity", size = 0) +
  scale_fill_manual(values = pal, 
                    name = "Nomenclature",
                    guide = guide_legend(nrow = 4,
                                         title.position = "top",
                                         title.hjust = 0 )) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 30, 5),
                     minor_breaks = seq(0, 30, 5),
                     limits = c(0, 35),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = "MOA") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        plot.margin = margin(t = 0, r = 0, l = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top",
        legend.text = element_text(size = 8, colour = "black"),
        legend.key.size = unit(0.1, "in")) +
  coord_flip()

# panel c
fig6c <- diffmoasum %>%
  ggplot(aes(x = MOA, y = n, fill = differentiation)) +
  geom_bar(stat = "identity", size = 0) +
   scale_fill_manual(values = c("#BABABA", pal[5]), 
                    name = "Differentiation capacity",
                    labels = c("Not demonstrated", "Demonstrated"),
                    guide = guide_legend(nrow = 4, 
                                         title.position = "top",
                                         title.hjust = 0)) +
  scale_y_continuous(name = "Number of Articles", 
                     breaks = seq(0, 30, 5),
                     minor_breaks = seq(0, 30, 5),
                     limits = c(0, 35),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0), name = NULL, labels = NULL) +
  theme_classic() +
  theme(axis.text.x = element_text(size = 8, colour = "black"),
        axis.text.y  = element_text(size = 8, colour = "black"),
        plot.margin = margin(t = 0, r = 0, l = 0),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "top",
        legend.text = element_text(size = 8, colour = "black"),
        legend.key.size = unit(0.1, "in")) +
  coord_flip()

@@

fig6 <- fig6a + fig6b + fig6c +
  plot_layout(design = layout) +
  plot_annotation(tag_levels = "A") &
    theme(plot.tag = element_text(size = 9, hjust = 0, face = "bold"))
ggsave("figs/fig_6.wmf", 
       plot = fig6, 
       device = "wmf",
       width = fig_w_2, 
       height = 3,
       units = units)

## relationship between phase_condensed and test_results

```{r}
res <- table(dat$phase_condensed, dat$test_results) %>%
  fisher.test(simulate.p.value = TRUE)
test_anno <- paste0("There is no significant association between MOA and cell categorisation\n(",
                   res$method, ": p = ", round(res$p.value, 3))

@@

# without the regen and multi cats
dat_sub <- dat %>%
  filter(stem_stromal != "Regenerative") %>% 
  filter(stem_stromal != "Multipotent Stromal") %>% droplevels()
res <- table(dat_sub$stem_stromal, dat_sub$MOA) %>% 
  fisher.test(simulate.p.value = TRUE)

@@

## Figure 10 assocation between differentiation and MOA

```{r}
res <- table(dat$differentiation, dat$MOA) %>% fisher.test(simulate.p.value = TRUE)

```
# CD values figures

@@

ggsave("figs/fig_x3_cd_refs_bar.wmf", 
       plot = cd_refs_bar, 
       device = "wmf",
       width = fig_w, 
       height = fig_w,
       units = units,
       dpi = dpi)

@@

```{r}
cd_values_point_isct <- cd_dat_summary_values %>% 
  filter(ISCT == "ISCT phenotype marker") %>% 
  ggplot(aes(x = marker, y = mean_value)) +
  geom_pointrange(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                  position = position_jitter(width = 0.5), shape = 20) +
  xlab("") +
  scale_y_continuous(name = "Value reported", 
                     breaks = seq(0, 110, 10),
                     minor_breaks = seq(0, 110, 10),
                     limits = c(0, 110),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure X4. TOP" ) +
  geom_vline(xintercept = seq(1.5, 10.5, 1), linetype = "longdash") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4))
ggsave("figs/fig_x4_cd_values_isct.wmf", 
       plot = cd_values_point_isct, 
       device = "wmf",
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)

@@

```{r}
cd_values_point_other <- cd_dat_summary_values %>% 
  filter(ISCT == "Other marker") %>% 
  ggplot(aes(x = marker, y = mean_value)) +
  geom_pointrange(aes(ymin = mean_value - se_value, ymax = mean_value + se_value), 
                  position = position_jitter(width = 0.5), shape = 20) +
  xlab("") +
  scale_y_continuous(name = "Value reported", 
                     breaks = seq(0, 110, 10),
                     minor_breaks = seq(0, 110, 10),
                     limits = c(0, 110),
                     expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  labs(title = "Wilson, Alison, Figure X4. TOP" ) +
  geom_vline(xintercept = seq(1.5, 10.5, 1), linetype = "longdash") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(size = 8, colour = "black"),
        axis.ticks.y = element_blank(),
        axis.text.y  = element_text(size = 8, colour = "black"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        # legend.title = element_text(size = 8, colour = "black"),
        # legend.text = element_text(size = 8, colour = "black"),
        # legend.key.size = unit(0.2, "in"),
        # legend.position = c(0.7, 0.8),
        legend.position = "none",
        plot.title = element_text(hjust = 0, size = 4))
ggsave("figs/fig_x5_cd_values_isct.wmf", 
       plot = cd_values_point_other, 
       device = "wmf",
       width = fig_w_1.5, 
       height = fig_w,
       units = units,
       dpi = dpi)
